---
title: 'MMGFM: simulation1'
author: "Wei Liu"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{MMGFM: simulation1}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

This vignette introduces the usage of MMGFM for the analysis of high-dimensional  multi-study multi-modality  data with additional covariates, by comparison with other methods. 


The package can be loaded with the command:
```{r  eval = FALSE}
library(MMGFM)
```

Next, we load some functions for subsequent use.
```{r  eval = FALSE}
source("https://raw.githubusercontent.com/feiyoung/MMGFM/refs/heads/main/simu_code/definedFunc.R")
```

## Generate the simulated data
First, we generate the simulated data from three data sources and three modalities, with each modality consisting of count variables. 
```{r  eval = FALSE}
N <- 100
q <- 3
qsvec  <- rep(2,3)
sigma_eps <- 1
datlist <- gendata_mmgfm(seed = 1,  nvec = c(300, 200, 100),  
                        pveclist = list('poisson'=c(50, 150, 200)),
                        q = q,  d= 3,qs = qsvec,  rho = 2, rho_z=0.5,
                        sigmavec=1,  sigma_eps=sigma_eps)
XList <- datlist$XList
max(unlist(XList))
print(str(XList))
ZList <- datlist$ZList # covariates
print(head(ZList[[1]]))
tauList <- datlist$tauList # offset term
numvarmat <- datlist$numvarmat
```


Fit the MMGFM model using the function `MMGFM()` in the R package `MMGFM`. Users can use `?MMGFM` to see the details about this function
```{r  eval = FALSE}
system.time({
  tic <- proc.time()
  reslist <- MMGFM(XList, ZList=ZList, numvarmat, q=q, qsvec = qsvec, init='MSFRVI')
  toc <- proc.time()
  time_MMGFM <- toc[3] - tic[3]
})
```


Check the increased property of the envidence lower bound function.
```{r  eval = FALSE}
library(ggplot2)
library(scales)
dat_iter <- data.frame(iter=1:length(reslist$ELBO_seq), ELBO=reslist$ELBO_seq)
ggplot(data=dat_iter, aes(x=iter, y=ELBO)) + geom_line() + geom_point() + theme_bw(base_size = 20) +
  scale_y_continuous(labels = label_scientific(digits = 8)) 

```

We calculate the metrics to measure the estimation accuracy, where the mean trace statistic is used to measure the estimation accuracy of loading matrix and prediction accuracy of factor matrix, which is evaluated by the function `measurefun()` in the R package `GFM`, and the root of mean absolute error is adopted to measure the estimation error of beta.
```{r  eval = FALSE}

methodNames <- c("MMGFM", "GFM", "MRRR", "MSFR", "MultiCOAP")
n_methods <- length(methodNames)
metricList <- list(F_tr = rep(NA, n_methods), 
                   H_tr = rep(NA,  n_methods), 
                   V_tr = rep(NA,  n_methods), 
                   A_tr = rep(NA, n_methods),
                   B_tr = rep(NA,  n_methods),
                   beta_norm=rep(NA,  n_methods),
                   time = rep(NA,  n_methods))
for(ii in seq_along(metricList)) names(metricList[[ii]]) <- methodNames
metricList$F_tr[1] <- meanTr(reslist$hF, datlist$F0List)
metricList$H_tr[1] <-meanTr(reslist$hH, datlist$H0List)
metricList$V_tr[1] <-meanTr(lapply(reslist$hv, function(x) Reduce(cbind,x) ), datlist$VList)
metricList$A_tr[1] <-metric_mean(AList=reslist$hA, datlist$A0List, align='unaligned', numvarmat = numvarmat)
metricList$B_tr[1] <- mean(ms_metric_mean(reslist$hB, datlist$B0List, align='unaligned', numvarmat = numvarmat))
metricList$beta_norm[1] <-normvec(Reduce(cbind, reslist$hbeta)- Reduce(cbind,datlist$betaList))
metricList$time[1] <- reslist$time.use

```


## Compare with other methods

We compare `MMGFM` with various prominent methods in the literature. They are (1)  Generalized factor model (Liu et al. 2023) implemented in the R package `GFM`; (2) Multi-response reduced-rank Poisson regression model (MMMR, Luo et al. 2018) implemented in `rrpack` R package; and (3) the multi-study covariate-augmented overdispersed Poisson factor (MultiCOAP) model.

(1). First, we implemented the generalized factor model (GFM) and record the metrics that measure the estimation accuracy and computational cost.

```{r  eval = FALSE}

res_gfm <- gfm_run(XList, numvarmat, q=q)
metricList$F_tr[2] <- meanTr(res_gfm$hF, datlist$F0List)
metricList$A_tr[2] <-metric_mean(AList=res_gfm$hA, datlist$A0List, align='unaligned', numvarmat = numvarmat)
metricList$time[2] <- res_gfm$time.use
```

(2). Then, we implemented Multi-response reduced-rank Poisson regression model (MMMR) and recorded the metrics. Here, we truncate values to ensure normal working of this function. Otherwise, it will produce error.
```{r  eval = FALSE}
res_mrrr <- mrrr_run(XList, ZList, numvarmat, q, truncflag=TRUE, trunc=500)
metricList$F_tr[3] <- meanTr(res_mrrr$hF, datlist$F0List)
metricList$A_tr[3] <- metric_mean(AList=res_mrrr$hA, datlist$A0List, align='unaligned', numvarmat = numvarmat)
metricList$beta_norm[3] <-normvec(res_mrrr$hbeta - Reduce(cbind,datlist$betaList))
metricList$time[3] <- res_mrrr$time.use
```

(3) Thirdly, we implemented the  zero-inflated Poisson factor model:
```{r  eval = FALSE}
source("https://raw.githubusercontent.com/feiyoung/MMGFM/refs/heads/main/simu_code/MSFR_main_R_MSFR_V1.R")
## To produce results in limited time, here we set maxIter=5. Even Set maxIter=1e4, the result is also not good.
res_msfr <- MSFR_run(XList, ZList, numvarmat, q, qs=qsvec, maxIter=5, load.source=TRUE, log.transform=TRUE)
metricList$F_tr[4] <- meanTr(res_msfr$hF, datlist$F0List)
metricList$H_tr[4] <- meanTr(res_msfr$hH, datlist$H0List)
metricList$A_tr[4] <- metric_mean(AList=res_msfr$hA, datlist$A0List, align='unaligned', numvarmat = numvarmat)
metricList$B_tr[4] <- mean(ms_metric_mean(res_msfr$hB, datlist$B0List, align='unaligned', numvarmat = numvarmat))
metricList$beta_norm[4] <- normvec(t(res_msfr$hbeta)- Reduce(cbind,datlist$betaList))
metricList$time[4] <- res_msfr$time.use
```
(4) Fourthly, we also applied the the multi-study covariate-augmented overdispersed Poisson factor (MultiCOAP) model.

```{r  eval = FALSE}
res_mcoap <- multicoap_run(XcList=XList, ZList,numvarmat, q, qsvec)
metricList$F_tr[5] <- meanTr(res_mcoap$hF, datlist$F0List)
metricList$H_tr[5] <-meanTr(res_mcoap$hH, datlist$H0List)
metricList$A_tr[5] <- metric_mean(AList=res_mcoap$hA, datlist$A0List, align='unaligned', numvarmat = numvarmat)
metricList$B_tr[5] <- mean(ms_metric_mean(res_mcoap$hB, datlist$B0List, align='unaligned', numvarmat = numvarmat))
metricList$beta_norm[5] <- normvec(t(res_mcoap$hbeta)- Reduce(cbind,datlist$betaList))
metricList$time[5] <- res_mcoap$time.use

```



## Comparison of performance

Next, we summarized the metrics for MMGFM and other compared methods in a data.frame object. We observed that MMGFM achieved much better estimation accuracy for the quantities of interest.

```{r  eval = FALSE}
mat.metric <- round(Reduce(rbind, metricList),3)
row.names(mat.metric) <- names(metricList)

dat_metric <- as.data.frame(mat.metric)
DT::datatable(dat_metric)
```

<details>
<summary>**Session Info**</summary>
```{r}
sessionInfo()
```
</details>
